---
title: "Ch_4"
author: "sbai200"
date: "07/04/2022"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# call methylaytion

## index bam files

```{bash, eval = FALSE}
data_folder=<file_path>

find $data_folder -name "*.bam" | xargs -I {} samtools index {} 
```

## install modbam2bed (v0.3.2)

```{bash, eval = FALSE}
conda create --name modbam2bed
source activate modbam2bed
conda install -c bioconda -c conda-forge -c epi2melabs modbam2bed
```

## generate bed file

```{bash, eval = FALSE}
data_folder=<file_path>
reference_genome=<file_path>

modbam2bed -e -m 5mC --cpg -t 4 =$reference_genome $data_folder/*.bam > guppy_female_meth_summary.cpg.bam
```

## save methjoin

```{python, eval = FALSE}

import pandas as pd

bedmethyl="guppy_female_meth_summary.cpg.bam"  #"guppy_female_genome_meth_summary.5mC.txt"
methdata = pd.read_csv(
bedmethyl, sep='\t',
header=None,
names=["chrom", "start", "end", "name", "score", "strand", "tstart", "tend", "color", "coverage", "freq", "canon", "mod", "filt"])

import aplanat
from aplanat import hist
from bokeh.layouts import gridplot
from aplanat import annot
from bokeh.plotting import output_file, save

names = ('fwd', 'rev')
fwdmeth = methdata.loc[methdata['strand'] == "+"]
revmeth = methdata.loc[methdata['strand'] == "-"]

# join the reverse to the fwd assuming sites are one apart
print("Joining forward and reverse strand joins assuming 1-base offset.")
tmp = revmeth.copy()
tmp['start'] -= 1
tmp['end'] -= 1
methjoin = pd.merge(fwdmeth, tmp, on=("chrom", "start"), suffixes=(".fwd", ".rev"))
methjoin["coverage"] = methjoin["coverage.fwd"] + methjoin["coverage.rev"]

#save data
methjoin.to_csv('guppy_female_methjoin_5mC.tsv', sep='\t')
```

## remove unused columns

```{bash, eval = FALSE}
cat guppy_female_genome_methjoin_5mC.tsv | cut -f 1,2,5,6,7,8 - > reduced_guppy_female_genome_methjoin_5mC.tsv
```

## mutate datasets

```{r}
library(readr)
library(dplyr)

methobj <- read_delim("./reduced_guppy_female_methjoin_5mC.tsv", delim = "\t")

# create new summaries female reads
methobj <- methobj %>% mutate(F_modbase = mod.fwd + mod.rev)
methobj <- methobj %>% mutate(F_freq = (F_modbase/coverage))
methobj <- methobj %>% mutate(F_percentmodbase = F_freq*100)

write.table(methobj, "mutated_reduced_guppy_female_methjoin_5mC.tsv", row.names = FALSE, sep = "\t")
```

## combine male and female

```{r}
library(readr)
library(dplyr)

options(warn = 1)

# column names
Female_colnames <- c("chrom", "start", "coverage", "F_modbase", "F_freq", "F_percentmodbase")
Male_colnames <- c("chrom", "start", "coverage", "M_modbase", "M_freq", "M_percentmodbase")

# load data
methobj <- read_delim("./mutated_reduced_guppy_female_methjoin_5mC.tsv", col_names = Female_colnames, delim = "\t", skip = 1)
methobj2 <- read_delim("./../Male/save_tables/mutated_reduced_guppy_male_methjoin_5mC.tsv", col_names = Male_colnames, delim = "\t", skip = 1)

# join only male and female dataset rows found in both sets
methobj3 <- inner_join(methobj, methobj2)

# create file with intersecting data
write.table(methobj3, "combined_5mC_InnerJoin.tsv", row.names = FALSE, col.names = TRUE, sep = "\t")
```

## correct the dataset to have all the female methylation calls for the W-linked contigs

```{r}
library(readr)
library(dplyr)

options(warn = 1)

# create list of col names
cname <- c("chrom", "start", "mod.fwd", "mod.rev", "coverage", "M_mod.fwd", "M_mod.rev", "M_coverage", "F_modbase", "F_freq", "F_percentmodbase", "M_modbase", "M_freq", "M_percentmodbase")

# load data
No_W.df <- read.table("./No_W_contigs_mutated_reduced_combined_reads_InnerJoin.tsv", col.names = cname, header = TRUE)
W.df <- read.table("./Edited_W_contigs_guppy_female_genome_methjoin.tsv", col.names = cname, header = TRUE)

methobj <- rbind(No_W.df, W.df)

write.table(methobj, "Corrected_mutated_reduced_combined_reads_InnerJoin.tsv", row.names = FALSE, sep = "\t")
```

# density plots

```{r}
library(readr)
library(dplyr)
library(ggplot2)
library(viridis)
library(magrittr)

options(warn = 1)

# create list of col names
cname <- c("chrom", "start", "mod.fwd", "mod.rev", "coverage", "M_mod.fwd", "M_mod.rev", "M_coverage", "F_modbase", "F_freq", "F_percentmodbase", "M_modbase", "M_freq", "M_percentmodbase")

# load data
methobj <- read.table("./Corrected_mutated_reduced_combined_reads_InnerJoin_201221.tsv", col.names = cname, header = TRUE)

# filter for sites with at least 4x coverage
methobj <- filter(methobj, coverage >= 5)
methobj <- filter(methobj, M_coverage >= 5)

# convert nas to zeros
methobj$F_percentmodbase[is.na(methobj$F_percentmodbase)] <- 0
methobj$M_percentmodbase[is.na(methobj$M_percentmodbase)] <- 0

# get contig lengths
fai.df <- read.table("./../female_curated_151121.fa.fai", col.names = c("chrom", "len", "index", "bases", "bytes"), header = FALSE)

# remove contigs smaller than 50 kb
fai.df %<>% filter(len >= 50000)
fai.df <- fai.df[,c(1,2)]
c <- fai.df[,1]

methobj <- methobj[methobj$chrom %in% c , ]

# import sex-linked contigs
sex_linked.df <- read_csv("./Z_or_W.csv")

# separate out Z- and W-linked contigs
Z.df <- sex_linked.df[1:27,1]
W.df <- sex_linked.df[,2]

# get list of sex-linked contig names
Z_vec <- pull(Z.df, Z)
W_vec <- pull(W.df, W)

# subset Z-linked contigs
Z_methobj.df <- methobj[methobj$chrom %in% Z_vec , ]
Z_methobj.df$Chromosome <- rep("Z-linked", nrow(Z_methobj.df))

# subset W-linked contigs
W_methobj.df <- methobj[methobj$chrom %in% W_vec , ]
W_methobj.df$Chromosome <- rep("W-linked", nrow(W_methobj.df))

# combine sex-linked contig IDS 
names(W.df)[names(W.df) == 'W'] <- 'Z'
sex_linked.df <- rbind(W.df, Z.df)
sex_linked_vec <- pull(sex_linked.df, Z)

# get list of all contig IDs
all_con <- pull(methobj, chrom)

# get autosomal and PAR contig IDs
A_con_vec <- setdiff(all_con, sex_linked_vec)

# create an autosomal and PAR subset
A_methobj.df <- methobj[methobj$chrom %in% A_con_vec , ]
A_methobj.df$Chromosome <- rep("Autosomal", nrow(A_methobj.df))

# generate two datasets, one with W-lnked contigs and one without
F_only_methobj <- rbind(Z_methobj.df, W_methobj.df, A_methobj.df)
Both_sexes_methobj <- rbind(Z_methobj.df, A_methobj.df)

F_only_methobj <- F_only_methobj %>% na.omit()
F_only_methobj$Chromosome <- as.factor(F_only_methobj$Chromosome)

Both_sexes_methobj <- Both_sexes_methobj %>% na.omit()
Both_sexes_methobj$Chromosome <- as.factor(Both_sexes_methobj$Chromosome)

# plot autosomal. PAR and Z-lnked contigs density
legend <- c("Female" = "#fc8d59", "Male"= "#0570b0")

all <- ggplot(data = Both_sexes_methobj) + geom_density(aes(M_percentmodbase, colour = "Male", fill = "Male"), alpha = .3) +
  geom_density(aes(F_percentmodbase, colour = "Female", fill = "Female"), alpha = .3) +
  theme_classic() +
  xlab("5mC %") +
  ylab("Density") +
  scale_color_manual(values = legend,name="Sex") +
  scale_fill_manual(values = legend,name="Sex") +
  theme(text = element_text(size = 24))

ggsave(plot = all,filename = "all_mean_methyl_50kb_plus.png", height=8.5, width=12.5)

# plot female percent 5mC CpG methylation for autosomal. PAR, Z- and W-linked contigs
Plots <- ggplot(data=F_only_methobj) + geom_density(aes(F_percentmodbase, colour = Chromosome, fill = Chromosome), alpha = .3) +
      scale_color_viridis(discrete=TRUE) +
      scale_fill_viridis(discrete = TRUE) +
      theme_classic() +
      xlab("5mC %") +
      ylab("Density") +
      guides(colour = guide_legend(title="Chromosome \ntype"),fill=guide_legend(title="Chromosome \ntype")) +
      theme(legend.position = c(.8, .85), text = element_text(size = 24))

ggsave(plot = Plots,filename = "female_mean_methyl_50kb_plus.png", height=8.5, width=12.5)

# get the mean and standard deviation of female 5mC CpG methylation for autosomal. PAR, Z- and W-linked contigs
T <- F_only_methobj %>% group_by(Chromosome) %>% summarise(ave = mean(F_percentmodbase), SD = sd(F_percentmodbase))
write.table(T, "over_50kb_compare_sets_summary_F2.txt", sep = "\t", row.names = FALSE)

# plot male percent 5mC CpG methylation for autosomal. PAR and Z-linked contigs
Plots <- ggplot(data=Both_sexes_methobj) + geom_density(aes(M_percentmodbase, colour = Chromosome, fill = Chromosome), alpha = .3) +
        scale_color_viridis(discrete=TRUE) +
        scale_fill_viridis(discrete = TRUE) +
        theme_classic() +
        xlab("5mC %") +
        ylab("Density") +
        guides(colour = guide_legend(title="Chromosome \ntype"),fill=guide_legend(title="Chromosome \ntype")) +
        theme(legend.position = c(.8, .85), text = element_text(size = 24))

ggsave(plot = Plots,filename = "male_mean_methyl_50kb_plus.png", height=8.5, width=12.5)

# get the mean and standard deviation of male 5mC CpG methylation for autosomal. PAR and Z-linked contigs
T <- Both_sexes_methobj %>% group_by(Chromosome) %>% summarise(ave = mean(M_percentmodbase), SD = sd(M_percentmodbase))
write.table(T, "over_50kb_compare_sets_summary_M2.txt", sep = "\t", row.names = FALSE)

# get the overall mean and standard deviation of male and female 5mC CpG methylation
mean(F_only_methobj$F_percentmodbase)
sd(F_only_methobj$F_percentmodbase)

mean(Both_sexes_methobj$F_percentmodbase)
sd(Both_sexes_methobj$F_percentmodbase)

mean(Both_sexes_methobj$M_percentmodbase)
sd(Both_sexes_methobj$M_percentmodbase)
```

# Contig end methylation and coverage

## classify dataset

```{r}
library(readr)
library(dplyr)
library(magrittr)

options(warn = 1)

# load data
df <- read.table("./../Both_sexes_corrected_mutated_reduced_combined_reads_InnerJoin_201221.tsv", header = TRUE)

# filter for sites with at least 5x coverage
df <- filter(df, coverage >= 5)
df <- filter(df, M_coverage >= 5)

# remove NAs
na.omit(df)

# low_F coverage sites
df_low_F <- df %>% filter(F_freq <=0.3)
df_low_F$F_classification <- rep("low_F", nrow(df_low_F))

# half coverage sites
df_intermediate_F <- df %>% filter(F_freq >0.3) %>% filter(F_freq <=0.7)
df_intermediate_F$F_classification <- rep("intermediate_F", nrow(df_intermediate_F))

# full coverage sites
df_high_F <- df %>% filter(F_freq >0.7) %>% filter(F_freq <=1)
df_high_F$F_classification <- rep("high_F", nrow(df_high_F))

# low_M coverage sites
df_low_M <- df %>% filter(M_freq <=0.3)
df_low_M$M_classification <- rep("low_M", nrow(df_low_M))

# half coverage sites
df_intermediate_M <- df %>% filter(M_freq >0.3) %>% filter(M_freq <=0.7)
df_intermediate_M$M_classification <- rep("intermediate_M", nrow(df_intermediate_M))

# full coverage sites
df_high_M <- df %>% filter(M_freq >0.7) %>% filter(M_freq <=1)
df_high_M$M_classification <- rep("high_M", nrow(df_high_M))

# join male and female subsets
f.df <- rbind(df_low_F, df_intermediate_F, df_high_F)
m.df <- rbind(df_low_M, df_intermediate_M, df_high_M)

# join datasets and combine male and female classifications
classified.df <- full_join(f.df, m.df)

# save
write.table(classified.df, "Both_sexes_classified_sites_whole_mutated_reduced_combined_reads_InnerJoin_180322.tsv", row.names = FALSE, sep = "\t")
```

## get level of methyation on ends

```{r}
library(readr)
library(dplyr)
library(magrittr)

options(warn = 1)

# load data
classified.df <- read.table("./classified_sites_whole_mutated_reduced_combined_reads_InnerJoin_180322.tsv", header = TRUE)

fai.df <- read.table("/nesi/nobackup/uoa02613/Hihi_MinION/FEMALE/Assembly_with_guppy_v5-0-7/curated_asm/female_split_assembly/female_asm.fai", col.names
 = c("ID", "length", "x1", "x2", "x3"))
fai.df <- fai.df[,1:2]

output <- matrix(data = NA, nrow=532, ncol=6)

# contig start

for (i in 1:nrow(fai.df)) {

  contig = fai.df[i,1]
  contig_name <- contig %>% unlist(use.names = FALSE)
  df1 <- filter(classified.df, chrom == contig_name)

  df2 <- df1 %>% filter(start <= 40000)

  mean_F <- mean(df2$F_freq)
  mean_M <- mean(df2$M_freq)

  median_F <- median(df2$F_freq)
  median_M <- median(df2$M_freq)

  output[i,1] <- contig_name
  output[i,2] <- "start"
  output[i,3] <- mean_F
  output[i,4] <- mean_M
  output[i,5] <- median_F
  output[i,6] <- median_M

}

output %<>% data.frame()
colnames(output) <- c("ID", "bin", "mean_F", "mean_M", "median_F", "median_M")

write.csv(output, "summary_contig_start_start_180322.csv", row.names = FALSE)

# contig ends

output <- matrix(data = NA, nrow=532, ncol=6)

for (i in 1:nrow(fai.df)) {
  end = fai.df[i,2] - 40000

  contig = fai.df[i,1]
  contig_name <- contig %>% unlist(use.names = FALSE)
  df1 <- filter(classified.df, chrom == contig_name)

  df3 <- df1 %>% filter(start >= end)

  mean_F <- mean(df3$F_freq)
  mean_M <- mean(df3$M_freq)

  median_F <- median(df2$F_freq)
  median_M <- median(df2$M_freq)

  output[i,1] <- contig_name
  output[i,2] <- "end"
  output[i,3] <- mean_F
  output[i,4] <- mean_M
  output[i,5] <- median_F
  output[i,6] <- median_M

}

output %<>% data.frame()
colnames(output) <- c("ID", "bin", "mean_F", "mean_M", "median_F", "median_M")

write.csv(output, "summary_contig_end_180322.csv", row.names = FALSE)

# interventions
fai.df <- read.csv("./interventions.csv", header = TRUE)

output <- matrix(data = NA, nrow=21, ncol=6)

# contig start
for (i in 1:nrow(fai.df)) {

  contig = fai.df[i,1]
  contig_name <- contig %>% unlist(use.names = FALSE)

  old_contig = fai.df[i,2]
  old_contig_name <- old_contig %>% unlist(use.names = FALSE)

  old_start = fai.df[i,3]
  old_end = fai.df[i,4]

  df1 <- classified.df %>% filter(chrom == contig_name)
  df1 <- filter(df1, start >= old_start)
  df1 <- filter(df1, start <= old_end)

  df2 <- df1 %>% filter(start <= 40000)

  mean_F <- mean(df2$F_freq)
  mean_M <- mean(df2$M_freq)

  median_F <- median(df2$F_freq)
  median_M <- median(df2$M_freq)

  output[i,1] <- contig_name
  output[i,2] <- old_contig_name
  output[i,3] <- mean_F
  output[i,4] <- mean_M
  output[i,5] <- median_F
  output[i,6] <- median_M

}

output %<>% data.frame()
colnames(output) <- c("ID", "bin", "mean_F", "mean_M", "median_F", "median_M")

write.csv(output, "summary_contig_start_interventions_180322.csv", row.names = FALSE)

# ends
output <- matrix(data = NA, nrow=21, ncol=6)

for (i in 1:nrow(fai.df)) {

  contig = fai.df[i,1]
  contig_name <- contig %>% unlist(use.names = FALSE)

  old_contig = fai.df[i,2]
  old_contig_name <- old_contig %>% unlist(use.names = FALSE)

  old_start = fai.df[i,3]
  old_end = fai.df[i,4]

  df1 <- classified.df %>% filter(chrom == contig_name)
  df1 <- filter(df1, start >= old_start)
  df1 <- filter(df1, start <= old_end)

  df2 <- df1 %>% filter(start <= 40000)

  mean_F <- mean(df2$F_freq)
  mean_M <- mean(df2$M_freq)

  median_F <- median(df2$F_freq)
  median_M <- median(df2$M_freq)

  output[i,1] <- contig_name
  output[i,2] <- old_contig_name
  output[i,3] <- mean_F
  output[i,4] <- mean_M
  output[i,5] <- median_F
  output[i,6] <- median_M

}

output %<>% data.frame()
colnames(output) <- c("ID", "bin", "mean_F", "mean_M", "median_F", "median_M")

write.csv(output, "summary_contig_end_interventions_180322.csv", row.names = FALSE)
```

## contig end cov

```{bash, eval = FALSE}
int=updated_female_consensus_renamed.fasta.fai.bed.sorted.40kbints.bed
genome=female_genome_edited.stat
ali1=Male_LR_female_asm
ali2=Female_LR_female_asm

samtools depth $ali1 | awk '{print $1"\t"$2"\t"$2+1"\t"$3}' | bedtools map -a  $int -b - -c 4 -o mean -g $genome > $ali1.40kbints_mean.txt

samtools depth $ali2 | awk '{print $1"\t"$2"\t"$2+1"\t"$3}' | bedtools map -a  $int -b - -c 4 -o mean -g $genome > $ali2.40kbints_mean.txt

# combine the interval data
cut -f 4 $ali2.40kbints_mean.txt | paste $ali1.40kbints_mean.txt - | awk '{print $1"\t"$2"\t"$3"\t"$4+$5}' > Combined_LR_female_asm.40kbints.mean.txt
```

## classifying contig end cov

```{r}
library(tidyverse)
library(magrittr)

#import interval means for 10kb and 40kb
mean_10kb.df <- read_delim("D:/Hihi_Project/Sarahs_scripts/New assembly/Contig end cov/Female/Combined_LR_female_asm.10kbints.mean.txt", col_names = c("Contig", "Start", "End", "Mean"))
mean_40kb.df <- read_delim("D:/Hihi_Project/Sarahs_scripts/New assembly/Contig end cov/Female/Combined_LR_female_asm.40kbints.mean.txt", col_names = c("Contig", "Start_40kb", "End_40kb", "Mean_40kb"))

#combine 10kb and 40kb means
mean_10_40kb_combined.df <- cbind(mean_10kb.df[,c(1,2,4)], mean_40kb.df[,c(2,4)])

#import interval medians for 10kb and 40kb
median_10kb.df <- read_delim("D:/Hihi_Project/Sarahs_scripts/New assembly/Contig end cov/Female/Combined_LR_female_asm.10kbints.median.txt", col_names = c("Contig", "Start", "End", "median"))
median_40kb.df <- read_delim("D:/Hihi_Project/Sarahs_scripts/New assembly/Contig end cov/Female/Combined_LR_female_asm.40kbints.median.txt", col_names = c("Contig", "Start_40kb", "End_40kb", "median_40kb"))

#combine 10kb and 40kb medians
median_10_40kb_combined.df <- cbind(median_10kb.df[,c(1,2,4)], median_40kb.df[,c(2,4)])
colnames(median_10_40kb_combined.df)

new.df <- cbind(mean_10_40kb_combined.df, median_10_40kb_combined.df[,c(3,5)])

#import per contig coverage (from misassemblies)
t <- list("c", "n", "n", "n", "n")
female_to_female_contig.df <- read_delim("D:/Hihi_Project/Sarahs_scripts/New assembly/Misassemblies/female_to_female_contig.coverage", delim = "\t", col_names = c("Contig", "Female_start", "Female_end", "Female_median", "Female_mean", "Female_count"), col_types = t, na = c("", "NA", "."))
male_to_female_contig.df <- read_delim("D:/Hihi_Project/Sarahs_scripts/New assembly/Misassemblies/male_to_female_contig.coverage", delim = "\t", col_names = c("Contig", "Male_start", "Male_end", "Male_median", "Male_mean", "Male_count"), col_types = t, na = c("", "NA", "."))
contig_M_F_combined.df <- cbind(female_to_female_contig.df[,c(1,4,5)], male_to_female_contig.df[,c(4,5)])

contig_M_F_combined.df %<>% mutate(combined_median = Female_median + Male_median, 
       combined_mean = Female_mean + Male_mean
       )

# combine mean, median interval and contig datasets
combined.df <- inner_join(new.df, contig_M_F_combined.df)
head(combined.df)

# get sets of chromosome types

W.df <- read_delim("D:/Hihi_Project/Sarahs_scripts/New assembly/Contig end cov/Female/female_W.txt", delim = "\t", col_names = "Contig") 

A.df <- read_delim("D:/Hihi_Project/Sarahs_scripts/New assembly/Contig end cov/Female/female_autosome.txt", delim = "\t", col_names = "Contig")

Z.df <- read_delim("D:/Hihi_Project/Sarahs_scripts/New assembly/Contig end cov/Female/female_Z.txt", delim = "\t", col_names = "Contig")

W_subset.df <- inner_join(combined.df, W.df)
Z_subset.df <- inner_join(combined.df, Z.df)
A_subset.df <- inner_join(combined.df, A.df)
```

## get the upper quartile of each set

```{r}
summary(W_subset.df[,c(12:13)])
```

```{r}
summary(Z_subset.df[,c(12:13)])
```

```{r}
summary(A_subset.df[,c(12:13)])
```

## use the upper quartile as the threshold for repetitive

```{r}
W_med_threshold <-12.00 
W_mean_threshold <- 31.847

Z_med_threshold <-47.00 
Z_mean_threshold <- 51.03

A_med_threshold <-57.00 
A_mean_threshold <- 57.53 
```

## label the comparison (does the interval seem repetitive Y/N?)

```{r}
W_subset.df$Median_repetitive10kb <- rep("NA", nrow(W_subset.df))
W_subset.df$Median_repetitive10kb[W_subset.df$median > W_med_threshold] <- "Y"
W_subset.df$Median_repetitive10kb[W_subset.df$median <= W_med_threshold] <- "N"

W_subset.df$Median_repetitive40kb <- rep("NA", nrow(W_subset.df))
W_subset.df$Median_repetitive40kb[W_subset.df$median_40kb > W_med_threshold] <- "Y"
W_subset.df$Median_repetitive40kb[W_subset.df$median_40kb <= W_med_threshold] <- "N"

W_subset.df$Mean_repetitive10kb <- rep("NA", nrow(W_subset.df))
W_subset.df$Mean_repetitive10kb[W_subset.df$Mean > W_mean_threshold] <- "Y"
W_subset.df$Mean_repetitive10kb[W_subset.df$Mean <= W_mean_threshold] <- "N"

W_subset.df$Mean_repetitive40kb <- rep("NA", nrow(W_subset.df))
W_subset.df$Mean_repetitive40kb[W_subset.df$Mean_40kb > W_mean_threshold] <- "Y"
W_subset.df$Mean_repetitive40kb[W_subset.df$Mean_40kb <= W_mean_threshold] <- "N"
head(W_subset.df)
```

## summarise the comparison

```{r}
W_subset.df %<>% unite("Median_summary", c("Median_repetitive10kb", "Median_repetitive40kb"), sep = "_")
W_subset.df %<>% unite("Mean_summary", c("Mean_repetitive10kb", "Mean_repetitive40kb"), sep = "_")
head(W_subset.df)
```

## classify the contig ends

```{r}
W_subset.df$repeat_size_median <- rep("NA", nrow(W_subset.df))
W_subset.df$repeat_size_median[W_subset.df$Median_summary == "Y_Y"] <- "long"
W_subset.df$repeat_size_median[W_subset.df$Median_summary == "N_Y"] <- "long"
W_subset.df$repeat_size_median[W_subset.df$Median_summary == "Y_N"] <- "short"
W_subset.df$repeat_size_median[W_subset.df$Median_summary == "N_N"] <- "not_repetitive"

W_subset.df$repeat_size_mean <- rep("NA", nrow(W_subset.df))
W_subset.df$repeat_size_mean[W_subset.df$Mean_summary == "Y_Y"] <- "long"
W_subset.df$repeat_size_mean[W_subset.df$Mean_summary == "N_Y"] <- "long"
W_subset.df$repeat_size_mean[W_subset.df$Mean_summary == "Y_N"] <- "short"
W_subset.df$repeat_size_mean[W_subset.df$Mean_summary == "N_N"] <- "not_repetitive"
head(W_subset.df)

write.csv(W_subset.df, "D:/Hihi_Project/Sarahs_scripts/New assembly/Contig end cov/Female/results/Classification_W_mean_and_median_100322.csv", row.names = FALSE)
```

## label the comparison (does the interval seem repetitive Y/N?)

```{r}
Z_subset.df$Median_repetitive10kb <- rep("NA", nrow(Z_subset.df))
Z_subset.df$Median_repetitive10kb[Z_subset.df$median > Z_med_threshold] <- "Y"
Z_subset.df$Median_repetitive10kb[Z_subset.df$median <= Z_med_threshold] <- "N"

Z_subset.df$Median_repetitive40kb <- rep("NA", nrow(Z_subset.df))
Z_subset.df$Median_repetitive40kb[Z_subset.df$median_40kb > Z_med_threshold] <- "Y"
Z_subset.df$Median_repetitive40kb[Z_subset.df$median_40kb <= Z_med_threshold] <- "N"

Z_subset.df$Mean_repetitive10kb <- rep("NA", nrow(Z_subset.df))
Z_subset.df$Mean_repetitive10kb[Z_subset.df$Mean > Z_mean_threshold] <- "Y"
Z_subset.df$Mean_repetitive10kb[Z_subset.df$Mean <= Z_mean_threshold] <- "N"

Z_subset.df$Mean_repetitive40kb <- rep("NA", nrow(Z_subset.df))
Z_subset.df$Mean_repetitive40kb[Z_subset.df$Mean_40kb > Z_mean_threshold] <- "Y"
Z_subset.df$Mean_repetitive40kb[Z_subset.df$Mean_40kb <= Z_mean_threshold] <- "N"
head(Z_subset.df)
```

## summarise the comparison

```{r}
Z_subset.df %<>% unite("Median_summary", c("Median_repetitive10kb", "Median_repetitive40kb"), sep = "_")
Z_subset.df %<>% unite("Mean_summary", c("Mean_repetitive10kb", "Mean_repetitive40kb"), sep = "_")
```

## classify the contig ends

```{r}
Z_subset.df$repeat_size_median <- rep("NA", nrow(Z_subset.df))
Z_subset.df$repeat_size_median[Z_subset.df$Median_summary == "Y_Y"] <- "long"
Z_subset.df$repeat_size_median[Z_subset.df$Median_summary == "N_Y"] <- "long"
Z_subset.df$repeat_size_median[Z_subset.df$Median_summary == "Y_N"] <- "short"
Z_subset.df$repeat_size_median[Z_subset.df$Median_summary == "N_N"] <- "not_repetitive"

Z_subset.df$repeat_size_mean <- rep("NA", nrow(Z_subset.df))
Z_subset.df$repeat_size_mean[Z_subset.df$Mean_summary == "Y_Y"] <- "long"
Z_subset.df$repeat_size_mean[Z_subset.df$Mean_summary == "N_Y"] <- "long"
Z_subset.df$repeat_size_mean[Z_subset.df$Mean_summary == "Y_N"] <- "short"
Z_subset.df$repeat_size_mean[Z_subset.df$Mean_summary == "N_N"] <- "not_repetitive"

write.csv(Z_subset.df, "D:/Hihi_Project/Sarahs_scripts/New assembly/Contig end cov/Female/results/Classification_Z_mean_and_median_100322.csv", row.names = FALSE)
```

## label the comparison (does the interval seem repetitive Y/N?)

```{r}
A_subset.df$Median_repetitive10kb <- rep("NA", nrow(A_subset.df))
A_subset.df$Median_repetitive10kb[A_subset.df$median > A_med_threshold] <- "Y"
A_subset.df$Median_repetitive10kb[A_subset.df$median <= A_med_threshold] <- "N"

A_subset.df$Median_repetitive40kb <- rep("NA", nrow(A_subset.df))
A_subset.df$Median_repetitive40kb[A_subset.df$median_40kb > A_med_threshold] <- "Y"
A_subset.df$Median_repetitive40kb[A_subset.df$median_40kb <= A_med_threshold] <- "N"

A_subset.df$Mean_repetitive10kb <- rep("NA", nrow(A_subset.df))
A_subset.df$Mean_repetitive10kb[A_subset.df$Mean > A_mean_threshold] <- "Y"
A_subset.df$Mean_repetitive10kb[A_subset.df$Mean <= A_mean_threshold] <- "N"

A_subset.df$Mean_repetitive40kb <- rep("NA", nrow(A_subset.df))
A_subset.df$Mean_repetitive40kb[A_subset.df$Mean_40kb > A_mean_threshold] <- "Y"
A_subset.df$Mean_repetitive40kb[A_subset.df$Mean_40kb <= A_mean_threshold] <- "N"
head(A_subset.df)
```

## summarise the comparison

```{r}
A_subset.df %<>% unite("Median_summary", c("Median_repetitive10kb", "Median_repetitive40kb"), sep = "_")
A_subset.df %<>% unite("Mean_summary", c("Mean_repetitive10kb", "Mean_repetitive40kb"), sep = "_")
head(A_subset.df)
```

## classify the contig ends

```{r}
A_subset.df$repeat_size_median <- rep("NA", nrow(A_subset.df))
A_subset.df$repeat_size_median[A_subset.df$Median_summary == "Y_Y"] <- "long"
A_subset.df$repeat_size_median[A_subset.df$Median_summary == "N_Y"] <- "long"
A_subset.df$repeat_size_median[A_subset.df$Median_summary == "Y_N"] <- "short"
A_subset.df$repeat_size_median[A_subset.df$Median_summary == "N_N"] <- "not_repetitive"

A_subset.df$repeat_size_mean <- rep("NA", nrow(A_subset.df))
A_subset.df$repeat_size_mean[A_subset.df$Mean_summary == "Y_Y"] <- "long"
A_subset.df$repeat_size_mean[A_subset.df$Mean_summary == "N_Y"] <- "long"
A_subset.df$repeat_size_mean[A_subset.df$Mean_summary == "Y_N"] <- "short"
A_subset.df$repeat_size_mean[A_subset.df$Mean_summary == "N_N"] <- "not_repetitive"

write.csv(A_subset.df, "D:/Hihi_Project/Sarahs_scripts/New assembly/Contig end cov/Female/results/Classification_autosomal_PAR_mean_and_median_100322.csv", row.names = FALSE)
```

Import methylaion data and coverage data into excel and combine with XLOOKUP().

## plot methylation and contig end cov

```{r}
library(readr)
library(magrittr)
library(dplyr)
library(ggplot2)
library(viridis)

dat <- read_csv("./../Results/Classifications_180322.csv", na = c("", "NA", "#N/A","#NUM!"))
head(dat)

dat$F_mean[is.na(dat$F_mean)] <- 0
dat$M_mean[is.na(dat$M_mean)] <- 0

dat[!complete.cases(dat), ]

dat <- dat[ , c(1,5,13,17:20)]

dat$repeat_size_mean <- as.factor(dat$repeat_size_mean)
levels(dat$repeat_size_mean) <- list("Long" = "long", "Not repetitive" = "not_repetitive", "Short" = "short")

dat$Set <- as.factor(dat$Set)
levels(dat$Set) <- list("Autosomal and PAR" = "A", "Z-linked" = "Z", "W-linked" = "W")


P <- ggplot(dat) +
  geom_point(aes(x = log10((F_mean*100) + 1), y = log10(Mean_40kb + 1), colour = repeat_size_mean), size = 4, alpha = .5) + 
  facet_grid(.~Set) + 
  scale_color_viridis(discrete=TRUE) +
  theme_bw() +
  theme(text =element_text(size = 20)) + 
  ylab("log10(mean coverage + 1)") + 
  xlab("log10(mean methylated sites + 1) [%]")
P <- P + guides(colour = guide_legend(title="Repeat \nclassification"))
P
#ggsave(plot = P,filename = "Methyl_end_cov_180322.png", height=8.5, width=17)
```

# high, medium and low methylated sites plot

```{r}

library(readr)
library(dplyr)

options(warn = 1)

# create list of col names
cname <- c("chrom", "start", "mod.fwd", "mod.rev", "coverage", "M_mod.fwd", "M_mod.rev", "M_coverage", "F_modbase", "F_freq", "F_percentmodbase", "M_modbase", "M_freq", "M_percentmodbase", "Chromosome", "F_classification", "M_classification")

# load data
methobj2 <- read_delim("./Both_sexes_classified_sites_whole_mutated_reduced_combined_reads_InnerJoin_180322.tsv", col_names = cname, delim = "\t", skip=1)

## FEMALE

# total number of observations

nrow(methobj2)

summary(methobj2$F_percentmodbase)
summary(methobj2$M_percentmodbase)

# low_F coverage sites
nrow(filter(methobj2, F_classification == "low_F"))

# half coverage sites
nrow(filter(methobj2, F_classification == "intermediate_F"))

# full coverage sites
nrow(filter(methobj2, F_classification == "high_F"))

## MALE

# low_M coverage sites
nrow(filter(methobj2, M_classification == "low_M"))

# half coverage sites
nrow(filter(methobj2, M_classification == "intermediate_M"))

# full coverage sites
nrow(filter(methobj2, M_classification == "high_M"))

```

## plot high medium and low sites

```{r}
library(readr)
library(ggplot2)
library(viridis)
library(reshape2)

whole.df <- read_delim("./../Data/whole_bins_180322.txt", delim = " ", comment = c("#"))


wholeF.df <- whole.df[2,c(1:4)]
colnames(wholeF.df) <- c("ID", "Low", "Intermediate", "High")
wholeF.df$sex <- "Female"

wholeM.df <- whole.df[2,c(1,5:7)]
colnames(wholeM.df) <- c("ID", "Low", "Intermediate", "High")
wholeM.df$sex <- "Male"

whole.df <- rbind(wholeF.df, wholeM.df)

data2 <- melt(whole.df, id.vars = "sex", measure.vars = c("Low", "Intermediate", "High"), 
              value.name="my_values", variable.name = "category")


#!!! CONFIGURE YOUR PLOT HERE !!!
# output
my_output <- paste("D:/Thesis/Methylation/plot_whole_figure.png",sep="/")
my_width <- 20
my_height <- 15
my_unit <- "cm"
 
# Colors colours <- c("Female methylation" = "#fc8d59", "Male methylation" = "#0570b0",
my_colors <- c("#fc8d59", "#0570b0")#, "#35608DFF")#,  "#1F988BFF")#c("#56B4E9", "#3492C7", "#F0E442", "#F04442")
# Bar height ratio
my_bar_height <- 0.75

# Font 
my_family <- "sans"
my_size_ratio <- 1.2

data2$sex <- factor(data2$sex)
#levels(data2$sex)

figure <- ggplot() + 
  geom_bar(aes(y = my_values, x = category, fill = sex), position = position_dodge(), data = data2, stat="identity", 
  width=my_bar_height) +
  theme_gray(base_size = 8) +
  scale_fill_manual(values = my_colors,labels =c("Female", "Male")) +
  xlab("\nLevel of Methylation") + 
  ylab("\nCount") +
  theme(legend.position="right",legend.title = element_blank()) + 
  theme(legend.text = element_text(family=my_family, size = rel(1.2)*my_size_ratio)) + 
  theme(panel.background = element_rect(color="#FFFFFF", fill="white")) + 
  theme(panel.grid.minor = element_blank()) + 
  theme(panel.grid.major = element_blank()) +
  theme(axis.text.y = element_text(family=my_family, colour = "black", size = rel(1.66)*my_size_ratio)) +
  theme(axis.text.x = element_text(family=my_family, colour = "black", size = rel(1.66)*my_size_ratio)) + 
  theme(axis.line = element_line(size=1*my_size_ratio, colour = "black")) + 
  theme(axis.ticks.length = unit(.85, "cm")) + 
  theme(axis.ticks.y = element_line(colour="white", size = 0)) + 
  theme(axis.ticks.x = element_line(colour="#222222")) + 
  theme(axis.ticks.length = unit(0.4, "cm")) + 
  theme(axis.title.y = element_text(family=my_family, size=rel(1.2)*my_size_ratio)) +
  theme(axis.title.x = element_text(family=my_family, size=rel(1.2)*my_size_ratio)) 

  ggsave(plot=figure, file=my_output, width = my_width, height = my_height, unit = my_unit)
figure
```

# correlation plot

```{r}
library(readr)
library(dplyr)
library(ggplot2)
library(rlang)
library(gridExtra)
library(ggpubr)
library(magrittr)

options(warn = 1)

# create list of col names
cname <- c("chrom", "start", "mod.fwd", "mod.rev", "coverage", "M_mod.fwd", "M_mod.rev", "M_coverage", "F_modbase", "F_freq", "F_percentmodbase", "M_mod
base", "M_freq", "M_percentmodbase")

# load data
methobj <- read.table("./No_W_contigs_mutated_reduced_combined_reads_InnerJoin_201221.tsv", col.names = cname, header = TRUE)

# filter for sites with at least 4x coverage
methobj <- filter(methobj, coverage >= 5)
methobj <- filter(methobj, M_coverage >= 5)

fai.df <- read.table("./../female_curated_151121.fa.fai", col.names = c("chrom", "len", "index", "bases", "bytes"), header = FALSE)

fai.df %<>% filter(len >= 50000)
fai.df <- fai.df[,c(1,2)]
c <- pull(fai.df, chrom)

methobj <- methobj[methobj$chrom %in% c , ]

res2 <-cor.test(methobj$F_percentmodbase, methobj$M_percentmodbase,  method = "pearson")
res2


P <- ggscatter(data = methobj, x = "M_percentmodbase", y = "F_percentmodbase",
        color = "black", alpha = .3,
        add = "reg.line",
        add.params = list(color = "blue", fill = "lightgray"),
        conf.int = TRUE,
        cor.coef = TRUE, cor.method = "pearson",
        xlab = "Female 5mC %", ylab = "Male 5mC %")


ggsave(plot = P, filename = "Over_50kb_pearson_cor_scatter.png", height = 8.5, width = 11)
```

