---
title: "Ch_3"
author: "sbai200"
date: "04/04/2022"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Combining homology-based and coverage-based evidence

## load packages
```{r}
library(tidyverse)
library(ggplot2)
library(magrittr)
```

## get list of haploid contigs

```{r}
haploid_types <- list("c", "i", "i", "i", "i", "n", "n", "n", "n", "i", "i", "n")

haploid_contigs.df <- read_csv("./LowCov alignments/50kb bins/Results/female_to_female_new_5kb_filter.csv", col_names = TRUE, col_types = haploid_types)
haploid_contigs.df <- haploid_contigs.df %>%  filter(Percent_hap > 0)
```

## nucmer alignments to Z and W

```{r}
nucmer_names <- c("S1", "E1", "S2", "E2", "LEN1", "LEN2", "%IDY", "Ref_ID", "Contig_id")
nucmer_types <- list("n","n","n","n","n","n","n","f","f")

# W Chromosome
Nucmer_W_ch_alignments.df <- read_delim("./Zebra finch alignments/Nucmer/Data/nucmer_female_medaka_to_mat_W_delta_filter - Copy.txt", delim = " ", col_names = nucmer_names, col_types = nucmer_types)

Nucmer_W_ch_alignments.df %<>% group_by(Contig_id) %>% summarise(Sum_len2 = sum(LEN2), Ave_identity = mean(`%IDY`)) 

# Z Chromosome
Nucmer_Z_ch_alignments.df <- read_delim("./Zebra finch alignments/Nucmer/Data/nucmer_female_medaka_to_pat_Z_delta_filter - Copy.txt", delim = " ", col_names = nucmer_names, col_types = nucmer_types)

Nucmer_Z_ch_alignments.df %<>% group_by(Contig_id) %>% summarise(Sum_len2 = sum(LEN2), Ave_identity = mean(`%IDY`))
```

## ragtag alignments to Z and W

```{r}
# W chromosome
RagTag_W_ch_alignments.df <- read_csv("./Zebra finch alignments/RagTag/Results/female_to_mat_W.csv", col_names = "Contig_id", skip = 1)

RagTag_Z_ch_alignments.df <- read_csv("./Zebra finch alignments/RagTag/Results/female_to_pat_Z.csv", col_names = "Contig_id", skip = 1)
```

## haploid contigs which intersect with ragtag and nucmer alignments to Z and W

```{r}
# W chromosome
Ragtag_and_nucmer_W_ch.df <- inner_join(RagTag_W_ch_alignments.df, Nucmer_W_ch_alignments.df)
haploid_contigs_in_ragtag_and_nucmer_W_ch.df <- inner_join(Ragtag_and_nucmer_W_ch.df, haploid_contigs.df)

write.table(haploid_contigs_in_ragtag_and_nucmer_W_ch.df, "haploid_contigs_in_ragtag_and_nucmer_W_ch.csv", sep = ",", col.names = TRUE, row.names = FALSE)

# Z chromosome
Ragtag_and_nucmer_Z_ch.df <- inner_join(RagTag_Z_ch_alignments.df, Nucmer_Z_ch_alignments.df)
haploid_contigs_in_ragtag_and_nucmer_Z_ch.df <- inner_join(Ragtag_and_nucmer_Z_ch.df, haploid_contigs.df)

write.table(haploid_contigs_in_ragtag_and_nucmer_Z_ch.df, "haploid_contigs_in_ragtag_and_nucmer_Z_ch.csv", sep = ",", col.names = TRUE, row.names = FALSE)
```

# Comparison of the male and female genomes

```{bash, eval=FALSE}

#!/bin/bash -e

#SBATCH --job-name=ragtag_female_medaka
#SBATCH --output=AW_%j.out
#SBATCH --error=AW_%j.err
#SBATCH --mail-user=sbai200@aucklanduni.ac.nz
#SBATCH --mail-type=END
#SBATCH --time=00:45:00
#SBATCH --mem=5G
#SBATCH --ntasks=1
#SBATCH --profile=task
#SBATCH --account=uoa00338

cd /nesi/nobackup/uoa02613/Hihi_MinION/Female/Downloads

module purge

module load Miniconda3

source activate ragtag2

echo "Starting to scaffold female assembly with male"
ragtag.py scaffold  -u Male_medaka_assembly Female_medaka_assembly
echo "Finished scaffolding"

echo "Moving files to their own directory"
mv ragtag_output/ female_medaka_to_male_medaka
echo "Done"
```

# Confirmation of sex-linked contigs

## load packages

```{r}
library(tidyverse)
library(ggplot2)
library(magrittr)
```

## plot M/F mapping ratio adapted from https://github.com/lurebgi/BOPsexChr/blob/master/m2f_ratio.r

```{r}
mappable_site_SR <- ggplot(sex_linked2.df) + geom_point(aes(y=SR_Female_mean,x=SR_m_f_map,size=Interval,fill=S_classification),shape=21) +
  ylab("Contig Mean (Female Coverage)")

mappable_site_SR <- mappable_site_SR +  scale_fill_discrete(name  ="Classification", 
                                                               breaks=c("Ch5", "W", "Z"), 
                                                               labels=c("Chromosome 5", "W-linked", "Z-linked"))

mappable_site_SR <- mappable_site_SR + scale_x_continuous(name = "Male/Female Mapping Ratio", limits = c(0,1))

mappable_site_SR <- mappable_site_SR + theme_classic()

mappable_site_SR <- mappable_site_SR +
    theme(text = element_text(size=12),
          legend.key.size = unit(1.5, "cm"),
          legend.key.width = unit(1.5,"cm") 
  )

my_output <- paste("D:/Hihi_Project/Sarahs_scripts/New assembly/Sex chromosome contig investigations/","mf_map_figure.png",sep="/")
my_width <- 20
my_height <- 15
my_unit <- "cm"

ggsave(mappable_site_SR, file=my_output, width = my_width, height = my_height, unit = my_unit)
```

# Identifying the pseudoautosomal region (PAR)

```{r}

```

